---
title: "Assignment 9 - K-means"
author: "Aytan Pirverdiyeva"  
format:
  html:
    code-fold: true 
    embed-resources: true
    toc: true
    number-sections: true
    theme: cerulean
---

# Goal of this assignment  
The goal of this assignment is for you to **explore the k-means algorithm** learned in class.  

In specific, you will be asked to:  
  - Use a different number of k than what we did in class,  
  - Validate the results of the new model, and  
  - Compare your results with those found in class  

# Instructions  
You will first need to watch (and code along with) the following k-means video:

  - k-means lab: https://youtu.be/GFgMp5tYiMU?si=CI0E-2r-cYZLGVL1 (start from beginning of video, watch till the 01:10)  
  - The partial script for the video above is in our GitHub repository, 04-classcode, **03-25_multivar_kmeans_partial.qmd**. Move the script mentioned above into the `code` subfolder of your `10_multivar` project
  
**AFTER watching the video and developing the code in class**, then start working on this HW assignment (instructions below):

  - Move this script into the `code` subfolder of your `10_multivar` project  
  - On this script, fill your name under the `title` section of the YAML  
  - Go back to the class k-means script, choose a different number of k than what we did in class, and run it.    

# Setup 
```{r}
#| message: false
#| warning: false

#install.packages("ggcorrplot")
#install.packages("factoextra")
#install.packages("ggpmisc")

library(tidyverse)
library(ggcorrplot)
library(broom)
library(car)
library(factoextra)
library(ggpmisc)
```


```{r}
weather <- read_csv("../../09_opendata/data/weather_monthsum.csv")

weather
```

```{r selecting only numerical vars}
weather_n <- weather %>%
  dplyr::select(-c(year:strength_gtex))

weather_n
```


```{r}
# normalizing the data
weather_norm <- weather_n%>%
  mutate(across(everything(), ~ scale(.x)))

weather_norm
summary(weather_norm)
```
```{r}
mod_km4 <- kmeans(weather_norm,
                 centers = 4,
                 nstart = 10
                 )

mod_km4
```
# Questions  
## 1. What number of k did you select? Why?
## For this experiment, I chose k=5 in order to test if an extra cluster will highlight any more nuanced patterns in the data that would have been clustered together in the 4-cluster model.

```{r kmeans model }
mod_km5 <- kmeans(weather_norm,
                 centers = 5,
                 nstart = 10
                 )

mod_km5

```


## 2. How many observations are there in each cluster?
```{r}
weather%>%
  mutate(cluster = mod_km5$cluster) %>%
  group_by(cluster) %>%
  tally()
```


## 3. Using the `fviz_cluster()` function, import here the original plot with k=4 did in class, and the new one with the number of k you selected for this exercise. How do they visually compare? Which one seems to be a better choice, and why?  

##Answer:
### Although there was a noticeable overlap between Cluster 1 (red) and Cluster 4 (purple) in the lower-middle region, the original figure with k = 4 demonstrated a respectable separation between the main clusters. This made interpretation a little more difficult, particularly in the vicinity of the center. 
###The k = 5 layout, on the other hand, divided that congested region into more precise groups, improving visual separation.  By establishing a unique cluster for the core region, the fifth cluster made it easier to separate patterns that had previously been grouped together.Despite the new cluster's tiny size (only 16 observations), it was simpler to understand the general structure and the distinctions across clusters.
###Since it minimizes overlap and captures more detail without being overly complicated, I think k = 5 provides a better clustering approach for this dataset.

```{r}
fviz_cluster(mod_km5,
             data = weather_norm)
```
```{r}
fviz_cluster(mod_km4,
             data = weather_norm)
```

# Submitting your work  
Once you have developed all the code and answers, make sure to Render this quarto file.  

**Notes on rendering**:  

- Make sure to render your work and inspect how the final html look like.  
- If it does not look professional for whatever reason, then fix the issue, re-render it, recheck.  
- Only send me your work once your html file looks professional.  
  - **DO NOT** delete the file's heading levels (# and ##). They set up the proper heading 1 and 2 levels, and I use them to guide my grading.  
  - If a given chunk is also outputting warnings or messages, inhibit this behavior by changing the chunk options `message` and `warning` to `FALSE` (I'll demo this in a moment).  
  
  - If, after rendered, 2 lines of text are connected and you wish to "break line" between them, add 2 extra spaces after the first one (I'lld demo this in a moment).  

After rendering, an .html file will be created on your `code` folder.  

Rename this file to `Assignment-09-kmeans-LASTNAME.html`.    

For ex., mine would be `Assignment-09-kmeans-Bastos.html`.

Submit your work by uploading the **html** file to **eLC Assignment #9 - K-means** by April 3rd 11:59 pm.  
  




